The Second Data Release of the INT Photometric H$\alpha$ Survey of the Northern Galactic Plane (IPHAS DR2)

### 2014-07-14 paper/iphas-dr2-rev1.tex
Added reliability subsection

>>>>

\subsection{Reliability}
\label{sec:reliability}

In producing a point source catalogue, spurious detections 
will inevitably be made. A fraction of these will mimic stellar 
sources closely enough that they cannot be excluded using the 
recorded parameters.

%The majority of sources have been observed in IPHAS more than 
%once (for objects with S/N $>$ 10 in the $r$-band, 62\% of 
%objects have secondary detections), however there are regions 
%observed only once - e.g. survey edges, intersections of inter-CCD 
%gaps. As such, requiring that objects have a secondary observation 
%to be considered reliable eliminates a significant fraction of
%genuine objects..

A statistic that can represent the confidence that objects in the 
catalogue are genuine is the reliability of the survey. The 
reliability gives the fraction of objects that can be repeatedly 
observed; provided the comparison is made to an external source of 
comparable depth this will provide the ratio of genuine astronomical 
sources to sources in the catalogue.

The IPHAS catalogue was compared to the SDSS DR7 point source 
catalogue (table \emph{PhotoObjAll}), in regions of 1 deg$^2$ 
chosen for contiguous coverage in both IPHAS and SDSS. This limited 
the choice of comparison regions to the second and third Galactic 
quadrants as the SDSS strips passing through the early plane are 
incomplete. DR7 was chosen for this comparison due to a change in 
the catalogue generation for DR8 which resulted in fewer stars being 
detected near bright ($r\,<\,10$) sources, which in turn gave an unduly 
pessimistic values for the reliability of IPHAS sources.

Comparisons were made for a number of subsets of the IPHAS point 
source catalogue: 
\begin{enumerate}
\item the entire (unfiltered) catalogue
\item objects detected in all three bands with \emph{mergedClass} equals -1, -2 or +1
\item $rErr\,<\,0.1$ and \emph{rClass} equals -1, -2 or +1
\item $iErr\,<\,0.1$ and \emph{iClass} equals -1, -2 or +1
\item \emph{reliable} sources
\item \emph{veryReliable} sources
\end{enumerate}
\noindent with the results presented in Table~\ref{tbl:reliability}. 
Requiring S/N $>$ 10 eliminates the majority of unreliable detections, 
as does selecting objects observed in multiple bands. The $i$-band performs 
marginally better than the $r$-band in terms of reliability, returning a 
slightly higher ($< 0.5\%$) reliability percentage for detections with S/N 
$>\,10$. Reliabilities $>99\%$ and $>99.5\%$ are achieved by using the 
supplied \emph{reliable} and \emph{veryReliable} quality flags.

The majority of unreliable sources in the IPHAS catalogue are 
located in the vicinity of bright ($r\,\lesssim\,12$) stars, tracing 
out the area surrounding the saturated source affected by scattered 
light. These can be eliminated in most cases by requiring detections 
in multiple bands.

\begin{table}
    \caption{Reliability values as computed by crossmatching the point 
             source catalogues of IPHAS and SDSS DR7 for a number of 
             subsets of IPHAS DR2 across 1 deg$^2$ regions.}
    \label{tbl:reliability}
    \begin{center}
        \begin{tabular}{lcccccc}
        \toprule
        Region ($\ell$, $b$) & Raw & nBands == 3 & $rErr\,<\,0.1$ & $iErr\,<\,0.1$ & \emph{reliable} & \emph{veryReliable} \\
        \midrule
        (149.39, 4.06) & 92.95 & 98.83 & 98.53 & 98.54 & 99.67 & 99.82 \\
        (186.59, -2.50) & 89.99 & 98.82 & 98.34 & 98.44 & 99.36 & 99.57 \\
        (202.70, -1.75) & 89.42 & 98.46 & 97.8 & 98.11 & 99.4 & 99.65 \\
        \bottomrule
        \end{tabular}
    \end{center}
\end{table}
